---
title: "Automatic Vs. Manual Transmission MPG"
author: "Emanuel Baisire"
output:
  word_document: default
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
---
### Executive Summary
The study purpose is to compare automatic and manual transmission engines'fuel consumption measured as Miles Per Gallon (MPG). The analysis is based from the 1974 Motor Trend US magazine "mtcars" dataset. A regressiona model is applied to determine better fuel efficiency transmission between manual and automatic transmission engines.

The hypothesis tested whether manual transmission engines are better than automatic in fuel consumption as measured by Miles Per Gallon (Mpg).A linear regression model results shows that manual transmission engines have a 7.2449 MPG advantage over autotamatic transmission.

### Cleaning data

#### Load and Test Data

The initial step was to load and categorize data from the Motor Trend magazine dataset. 
```{r}
data(mtcars)
 str(mtcars)
```

### Process Data
To proceed,it was necessary to convert some numeric variables to factors for a better data analysis.

```{r}
mtcars$cyl <- factor(mtcars$cyl)
mtcars$vs <- factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am,labels=c('Automatic','Manual'))
```

###: Exploratory analysis
This is helpful in understanding the data and the main variables to be considered for the model creation.

```{r}
summary(mtcars)
```

### Creating a Regression Model:  

A simple Linear Regression and a Multivariate Regression were used and compared against each other to determine the best results. ANOVA was also applied for the 2 models to determine if there was any strong variances.

```{r}
fit <- lm(mpg~am, data = mtcars)
summary(fit)

best_fit <- lm(mpg~am + wt + hp, data = mtcars)
anova(fit, best_fit)
```
### Model Testing:

```{r}
t.test(mpg ~ am, data = mtcars)
```
### Boxplot
```{r}
boxplot(mpg~am, data = mtcars,col ="blue",
xlab = "Transmission",
ylab = "Miles per Gallon",
main = "MPG by Transmission Type")
```


###Results: 
The boxplot indicates that a manual transmission engine achieves extra miles per gallon compared to an automatic transmission.

The t-test output of a p-value: 0.001374 that is less than 0.05 strongly validates our results.

### Appendix: 1 
Cars-Scatterplot:
```{r}
pairs(mpg ~ ., data = mtcars)
```

### Appendix: 2
MPG Boxplots
```{r}
par(mfrow=c(1,1))
with(mtcars,{
     boxplot(mpg~am, data = mtcars,col ="blue",
xlab = "Transmission",
ylab = "Miles per Gallon",
main = "MPG by Transmission Type")
     plot(mpg ~ as.integer(am),
             xlab = "Automatic (1) or Manual(2)",
             ylab = "miles per gallon (MPG)")
     abline(fit, col=4)
     })
```
